---
title: "Lab 5"
author: "Sarah Lam"
date: "10/26/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
    number_sections: yes
    code_folding: hide #inserts code button in document so that reader has the choice to view code where applicable 
---

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Confidence Intervals 

In this section we'll learn to use the 't.test()' function as a way to return the confidence interval using the t-distribution.

## Example 1 

Create a mock sample of observation values, and find that confidence interval. Then we'll learn how to report that. 

```{r}
pika_mass <- c(82, 94, 110, 700, 86, 99, 102, 320, 5000) # first question is to ask "is this representative of my population?"

pika_mean <- mean(pika_mass)

pika_sd <- sd(pika_mass)

pika_ci <- t.test(pika_mass)
```

### Risky & tedious way: 

Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g (mean $\pm$ 1 standard deviation, n= 7) with a 95% confidence interval of [79.4, 104.4] g.

### High initial investment, but safer & better way 

In-line code referencing to stored objects!

Use a single backtick on either side of a lowercase r to create an in-line bit of code. 

Mean pika mass at Paiute Pass is `r round(pika_mean, 1)`g $\pm$ `r round(pika_sd, 1)`g (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)`) with a 95% confidence interval of [`r round(pika_ci$conf.int[1], 1)`, `r round(pika_ci$conf.int[2], 1)`] g. 

# Two-sample t-test

Use a two sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference in means = 0). 

- H0: Means difference = 0 
- HA: Means difference is NOT = 0 

This is a two-sided t-test because no directionality is implied. 

## Example 1 

```{r}

oak_trees <- c(29, 19, 22, 30 , 35, 16, 20, 7)

pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19, 20)
```

Is there a significant difference in mean height for oak and pine trees? 

```{r}
trees_t <- t.test(oak_trees, pine_trees)
```

The p-value of `r trees_t$p.value` means that there is a `r (trees_t$p.value)*100`% chance that if the samples were drawn from populations with the same mean, we would get sample means this different. 
alt - there is a 1.8% chance of finding sample means *at least as different as those I found* by random chance if they were drawn from populations with the same height. 

## Example 2 

Using the 'mpg' dataset to compare city gas mileage to SUVs and compact cars. 

```{r}
# Create a subset called compact_suv that only contains observations where the class is suz OR compact
compact_suv <- mpg %>% 
  filter(class %in% c("suv", "compact"))
```

Side by side histograms and side by side QQ-plots to explore normality 
```{r}
ggplot(data = compact_suv, aes(x = cty)) +
  geom_histogram(bins = 12) +
  facet_wrap(~ class)

ggplot(data = compact_suv, aes(sample = cty)) +
  geom_qq() +
  facet_wrap(~ class)
```

```{r}
car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarize(mean_city = mean(cty), 
            sd_city = sd(cty), 
            sample_size = n())

car_stats
```
```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty)

suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)

cars_t <- t.test(compact_sample, suv_sample)
```

stats speak: I would reject the null hypothesis of equal mean city gas mileages 

Means gas mileage for compact cars and SUVs differed significantly (t(`r cars_t$parameter`) = `r cars_t$statistic`, p < 0.001. $\alpha$ = 0.05). 
